---
title: "Family and Schools Together <br /> Measurement Models"
author: "Jay Jeffries"
date: "11/13/2022"
output:
  rmdformats::robobook:
    code_folding: show
    self_contained: true
    thumbnails: false
    lightbox: true
pkgdown:
  as_is: true    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)

p_load("tidyverse", "lavaan", "skimr", "haven", "semPlot", "tidySEM", "rmdformats", "semTable", "htmlTable",
       "patchwork",
       install = T)

FASTData <- read_csv("/Users/jayjeffries/Desktop/Longitudinal/SEM/Assignments/Final Project/FAST SEM Project/FASTData.csv")
```

# Initial Measurement Model
## Model Syntax

```{r, warning = F}
meas_mod <- '
Soc_Cap =~ Q8A + Q8B + Q8C + Q9A + Q9B + Q9C + Q11 + Q10

Soc_Cap ~~ 1*Soc_Cap
'

fit_meas <- sem(model = meas_mod, data = FASTData, estimator = "ML", missing = "ML")
results <- summary(fit_meas, fit.measures = T)
results
```

```{r, warning = F, echo = F}
meas_mod_pre <- '
SC_pre =~ NA*P6 + lambda1*P6 + lambda2*P5 + lambda3*P3A + lambda4*P3B + lambda5*P3C + lambda6*P4A + lambda7*P4B + lambda8*P4C

SC_pre ~~ SC_pre
SC_pre ~ 1

P6 ~ int1*1
P5 ~ int2*1
P3A ~ int3*1
P3B ~ int4*1 
P3C ~ int5*1
P4A ~ int6*1
P4B ~ int7*1 
P4C ~ int8*1

0 == (int1 + int2 + int3 + int4 + int5 + int6 + int7 + int8)/8
1 == (lambda1 + lambda2 + lambda3 + lambda4 + lambda5 + lambda6 + lambda7 + lambda8)/8

# Covariances between relational trust items
P3A ~~ P3B + P3C + P4A + P4B + P4C 
P3B ~~ P3C + P4A + P4B + P4C 
P3C ~~ P4A + P4B + P4C 
P4A ~~ P4B + P4C 
P4B ~~ P4C
'
#fit_meas_pre <- sem(model = meas_mod_pre, data = FASTData, estimator = "ML", missing = "ML")
#summary(fit_meas_pre, fit.measures = T)

meas_mod_y1 <- '
SC_y1 =~ NA*Q11 + lambda9*Q11 + lambda10*Q10 + lambda11*Q8A + lambda12*Q8B + lambda13*Q8C + lambda14*Q9A + lambda15*Q9B + lambda16*Q9C

SC_y1 ~~ SC_y1
SC_y1 ~ 1

Q8A ~ int9*1
Q8B ~ int10*1
Q8C ~ int11*1
Q9A ~ int12*1 
Q9B ~ int13*1
Q9C ~ int14*1
Q11 ~ int15*1 
Q10 ~ int16*1
0 == (int9 + int10 + int11 + int12 + int13 + int14 + int15 + int16)/8
1 == (lambda9 + lambda10 + lambda11 + lambda12 + lambda13 + lambda14 + lambda15 + lambda16)/8

# Covariances between relational trust items
Q8A ~~ Q8B + Q8C + Q9A + Q9B + Q9C 
Q8B ~~ Q8C + Q9A + Q9B + Q9C 
Q8C ~~ Q9A + Q9B + Q9C 
Q9A ~~ Q9B + Q9C 
Q9B ~~ Q9C
'
#fit_meas_y1 <- sem(model = meas_mod_y1, data = FASTData, estimator = "ML", missing = "ML")
#summary(fit_meas_y1, fit.measures = T)

meas_mod_y2 <- '
SC_y2 =~ NA*R6 + lambda17*R6 + lambda18*R5 + lambda19*R3A + lambda20*R3B + lambda21*R3C + lambda22*R4A + lambda23*R4B + lambda24*R4C

SC_y2 ~~ SC_y2
SC_y2 ~ 1

R6 ~ int17*1
R5 ~ int18*1
R3A ~ int19*1
R3B ~ int20*1 
R3C ~ int21*1
R4A ~ int22*1
R4B ~ int23*1 
R4C ~ int24*1
0 == (int17 + int18 + int19 + int20 + int21 + int22 + int23 + int24)/8
1 == (lambda17 + lambda18 + lambda19 + lambda20 + lambda21 + lambda22 + lambda23 + lambda24)/8

# Covariances between relational trust items
R3A ~~ R3B + R3C + R4A + R4B + R4C 
R3B ~~ R3C + R4A + R4B + R4C 
R3C ~~ R4A + R4B + R4C 
R4A ~~ R4B + R4C 
R4B ~~ R4C
'
#fit_meas_y2 <- sem(model = meas_mod_y2, data = FASTData, estimator = "ML", missing = "ML")
#summary(fit_meas_y2, fit.measures = T)

meas_mod_y3 <- '
SC_y3 =~ NA*S12A + lambda25*S12A + lambda26*S11 + lambda27*S9A + lambda28*S9B + lambda29*S9C + lambda30*S10A + lambda31*S10B + lambda32*S10C

SC_y3 ~~ SC_y3
SC_y3 ~ 1

S12A ~ int25*1
S11 ~ int26*1
S9A ~ int27*1
S9B ~ int28*1 
S9C ~ int29*1
S10A ~ int30*1
S10B ~ int31*1 
S10C ~ int32*1
0 == (int25 + int26 + int27 + int28 + int29 + int30 + int31 + int32)/8
1 == (lambda25 + lambda26 + lambda27 + lambda28 + lambda29 + lambda30 + lambda31 + lambda32)/8

# Covariances between relational trust items
S9A ~~ S9B + S9C + S10A + S10B + S10C 
S9B ~~ S9C + S10A + S10B + S10C 
S9C ~~ S10A + S10B + S10C 
S10A ~~ S10B + S10C 
S10B ~~ S10C
'

fit_measure_mod_pre <- sem(model = meas_mod_pre, data = FASTData, estimator = "MLR", missing = "ML")
fit_measure_mod_y1 <- sem(model = meas_mod_y1, data = FASTData, estimator = "MLR", missing = "ML")
fit_measure_mod_y2 <- sem(model = meas_mod_y2, data = FASTData, estimator = "MLR", missing = "ML")
fit_measure_mod_y3 <- sem(model = meas_mod_y3, data = FASTData, estimator = "MLR", missing = "ML")

semTable(fit_measure_mod_pre, columns = c("est", "se", "z", "p", "rsquare"), type = "html", print.results = F, 
         fits = c("npar", "chisq", "df", "pvalue", "chisq.scaled", "df.scaled", "pvalue.scaled", "cfi", 
                  "rmsea", "srmr")) |>
  htmlTable()
semTable(fit_measure_mod_y1, columns = c("est", "se", "z", "p", "rsquare"), type = "html", print.results = F, 
         fits = c("npar", "chisq", "df", "pvalue", "chisq.scaled", "df.scaled", "pvalue.scaled", "cfi", 
                  "rmsea", "srmr")) |>
  htmlTable()
semTable(fit_measure_mod_y2, columns = c("est", "se", "z", "p", "rsquare"), type = "html", print.results = F, 
         fits = c("npar", "chisq", "df", "pvalue", "chisq.scaled", "df.scaled", "pvalue.scaled", "cfi", 
                  "rmsea", "srmr")) |>
  htmlTable()
semTable(fit_measure_mod_y3, columns = c("est", "se", "z", "p", "rsquare"), type = "html", print.results = F, 
         fits = c("npar", "chisq", "df", "pvalue", "chisq.scaled", "df.scaled", "pvalue.scaled", "cfi", 
                  "rmsea", "srmr")) |>
  htmlTable()
```

## Mod Indices

```{r}
modindices(fit_meas, sort = T)
```

## Measurement Model Figure

```{r, echo = F}
prepare_graph(
  edges = get_edges(fit_measure_mod_y1) |> filter(!str_detect(label_results, "WITH")),
  nodes = get_nodes(fit_measure_mod_y1) |> mutate(label = name), 
  layout = get_layout("", "", "", "", "SC_y1", "","", "",
                      "Q8A", "Q8B", "Q8C", "Q9A", "Q9B", "Q9C", "Q11", "Q10", rows = 2), 
  spacing_y = 2, spacing_x = 1.6, text_size = 1.7, angle = 170, 
  ellipse_width = 5, variance_diameter = .25) |> 
  edit_graph({label = paste(est_sig_std, "\n", confint_std) }) |>
  edit_graph({label = paste(name, "\n", est_sig_std, "\n", confint_std) }, element = "nodes") |>
  plot()
```

## Tau-Equivalence Test

```{r, warning = F}
congeneric_mod <- '
Soc_Cap =~ NA*Q8A + Q8B + Q8C + Q9A + Q9B + Q9C + Q11 + Q10

Soc_Cap ~~ 1*Soc_Cap

# Covariances between relational trust items
Q8A ~~ Q8B + Q8C + Q9A + Q9B + Q9C 
Q8B ~~ Q8C + Q9A + Q9B + Q9C 
Q8C ~~ Q9A + Q9B + Q9C 
Q9A ~~ Q9B + Q9C 
Q9B ~~ Q9C 
'

tau_mod <- '
Soc_Cap =~ equal*Q8A + equal*Q8B + equal*Q8C + equal*Q9A + equal*Q9B + equal*Q9C + equal*Q11 + equal*Q10

Soc_Cap ~~ 1*Soc_Cap

# Covariances between relational trust items
Q8A ~~ Q8B + Q8C + Q9A + Q9B + Q9C 
Q8B ~~ Q8C + Q9A + Q9B + Q9C 
Q8C ~~ Q9A + Q9B + Q9C 
Q9A ~~ Q9B + Q9C 
Q9B ~~ Q9C 
'
```

```{r, include = F}
fit_cong <- sem(model = congeneric_mod, data = FASTData, estimator = "ML", missing = "ML")
fit_tau <- sem(model = tau_mod, data = FASTData, estimator = "ML", missing = "ML")

summary(fit_cong, fit.measures = T)
summary(fit_tau, fit.measures = T)
```

```{r}
anova(fit_cong, fit_tau)
```

## Tau-Equivalence Results

The significant difference in model, where the tau-equivalent model fit worse in comparison to the congeneric model ($\chi^2_{tau}(13)=781.953$, $\chi^2_{cong}(5)=18.191$, $\triangle\chi^2=745.762$, $\triangle df=8,  p < .001$), indicates that a latent variable would be preferable to aggregated scores. This also means that a key assumption of Cronbach’s alpha is not met.

# Measurement Invariance

## Configural Invariance Test

```{r, warning = F, echo = F}
covs <- '
P6 ~~ Q11 + R6 + S12A
Q11 ~~ R6 + S12A
R6 ~~ S12A

P5 ~~ Q10 + R5 + S11
Q10 ~~ R5 + S11
R5 ~~ S11

P3A ~~ Q8A + R3A + S9A
Q8A ~~ R3A + S9A
R3A ~~ S9A

P3B ~~ Q8B + R3B + S9B
Q8B ~~ R3B + S9B
R3B ~~ S9B

P3C ~~ Q8C + R3C + S9C
Q8C ~~ R3C + S9C
R3C ~~ S9C

P4A ~~ Q9A + R4A + S10A
Q9A ~~ R4A + S10A
R4A ~~ S10A

P4B ~~ Q9B + R4B + S10B
Q9B ~~ R4B + S10B
R4B ~~ S10B

P4C ~~ Q9C + R4C + S10C
Q9C ~~ R4C + S10C
R4C ~~ S10C

SC_pre ~~ SC_y1 + SC_y2 + SC_y3
SC_y1 ~~ SC_y2 + SC_y3
SC_y2 ~~ SC_y3
'
#fit_meas_y3 <- sem(model = meas_mod_y3, data = FASTData, estimator = "ML", missing = "ML")
#summary(fit_meas, fit.measures = T)

configural <- paste(meas_mod_pre, meas_mod_y1, meas_mod_y2, meas_mod_y3, covs)
```

```{r, warning = F}
fit_configural <- sem(model = configural, data = FASTData, estimator = "ML", missing = "ML")
fitMeasures(fit_configural, fit.measures = c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
```

## Weak Invariance Model
```{r, warning = F}
weak_meas_mod <- '
SC_pre =~ NA*P6 + lambda1*P6 + lambda2*P5 + lambda3*P3A + lambda4*P3B + lambda5*P3C + lambda6*P4A + lambda7*P4B + lambda8*P4C

SC_pre ~~ SC_pre
SC_pre ~ 1

P6 ~ int1*1
P5 ~ int2*1
P3A ~ int3*1
P3B ~ int4*1 
P3C ~ int5*1
P4A ~ int6*1
P4B ~ int7*1 
P4C ~ int8*1

0 == (int1 + int2 + int3 + int4 + int5 + int6 + int7 + int8)/8
1 == (lambda1 + lambda2 + lambda3 + lambda4 + lambda5 + lambda6 + lambda7 + lambda8)/8

# Covariances between relational trust items
P3A ~~ P3B + P3C + P4A + P4B + P4C 
P3B ~~ P3C + P4A + P4B + P4C 
P3C ~~ P4A + P4B + P4C 
P4A ~~ P4B + P4C 
P4B ~~ P4C

SC_y1 =~ NA*Q11 + lambda1*Q11 + lambda2*Q10 + lambda3*Q8A + lambda4*Q8B + lambda5*Q8C + lambda6*Q9A + lambda7*Q9B + lambda8*Q9C

SC_y1 ~~ SC_y1
SC_y1 ~ 1

Q8A ~ int9*1
Q8B ~ int10*1
Q8C ~ int11*1
Q9A ~ int12*1 
Q9B ~ int13*1
Q9C ~ int14*1
Q11 ~ int15*1 
Q10 ~ int16*1
0 == (int9 + int10 + int11 + int12 + int13 + int14 + int15 + int16)/8

# Covariances between relational trust items
Q8A ~~ Q8B + Q8C + Q9A + Q9B + Q9C 
Q8B ~~ Q8C + Q9A + Q9B + Q9C 
Q8C ~~ Q9A + Q9B + Q9C 
Q9A ~~ Q9B + Q9C 
Q9B ~~ Q9C

SC_y2 =~ NA*R6 + lambda1*R6 + lambda2*R5 + lambda3*R3A + lambda4*R3B + lambda5*R3C + lambda6*R4A + lambda7*R4B + lambda8*R4C

SC_y2 ~~ SC_y2
SC_y2 ~ 1

R6 ~ int17*1
R5 ~ int18*1
R3A ~ int19*1
R3B ~ int20*1 
R3C ~ int21*1
R4A ~ int22*1
R4B ~ int23*1 
R4C ~ int24*1
0 == (int17 + int18 + int19 + int20 + int21 + int22 + int23 + int24)/8

# Covariances between relational trust items
R3A ~~ R3B + R3C + R4A + R4B + R4C 
R3B ~~ R3C + R4A + R4B + R4C 
R3C ~~ R4A + R4B + R4C 
R4A ~~ R4B + R4C 
R4B ~~ R4C

SC_y3 =~ NA*S12A + lambda1*S12A + lambda2*S11 + lambda3*S9A + lambda4*S9B + lambda5*S9C + lambda6*S10A + lambda7*S10B + lambda8*S10C

SC_y3 ~~ SC_y3
SC_y3 ~ 1

S12A ~ int25*1
S11 ~ int26*1
S9A ~ int27*1
S9B ~ int28*1 
S9C ~ int29*1
S10A ~ int30*1
S10B ~ int31*1 
S10C ~ int32*1
0 == (int25 + int26 + int27 + int28 + int29 + int30 + int31 + int32)/8

# Covariances between relational trust items
S9A ~~ S9B + S9C + S10A + S10B + S10C 
S9B ~~ S9C + S10A + S10B + S10C 
S9C ~~ S10A + S10B + S10C 
S10A ~~ S10B + S10C 
S10B ~~ S10C
'

covs <- '
P6 ~~ Q11 + R6 + S12A
Q11 ~~ R6 + S12A
R6 ~~ S12A

P5 ~~ Q10 + R5 + S11
Q10 ~~ R5 + S11
R5 ~~ S11

P3A ~~ Q8A + R3A + S9A
Q8A ~~ R3A + S9A
R3A ~~ S9A

P3B ~~ Q8B + R3B + S9B
Q8B ~~ R3B + S9B
R3B ~~ S9B

P3C ~~ Q8C + R3C + S9C
Q8C ~~ R3C + S9C
R3C ~~ S9C

P4A ~~ Q9A + R4A + S10A
Q9A ~~ R4A + S10A
R4A ~~ S10A

P4B ~~ Q9B + R4B + S10B
Q9B ~~ R4B + S10B
R4B ~~ S10B

P4C ~~ Q9C + R4C + S10C
Q9C ~~ R4C + S10C
R4C ~~ S10C

SC_pre ~~ SC_y1 + SC_y2 + SC_y3
SC_y1 ~~ SC_y2 + SC_y3
SC_y2 ~~ SC_y3
'
#fit_meas_y3 <- sem(model = meas_mod_y3, data = FASTData, estimator = "ML", missing = "ML")
#summary(fit_meas, fit.measures = T)

weak <- paste(weak_meas_mod, covs)
```

```{r, warning = F}
fit_weak <- sem(model = weak, data = FASTData, estimator = "ML", missing = "ML")

fitMeasures(fit_weak, fit.measures = c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
```

## Strong Invariance Model
```{r, warning = F}
strong_meas_mod <- '
SC_pre =~ NA*P6 + lambda1*P6 + lambda2*P5 + lambda3*P3A + lambda4*P3B + lambda5*P3C + lambda6*P4A + lambda7*P4B + lambda8*P4C

SC_pre ~~ SC_pre
SC_pre ~ 1

P6 ~ int1*1
P5 ~ int2*1
P3A ~ int3*1
P3B ~ int4*1 
P3C ~ int5*1
P4A ~ int6*1
P4B ~ int7*1 
P4C ~ int8*1

0 == (int1 + int2 + int3 + int4 + int5 + int6 + int7 + int8)/8

# Covariances between relational trust items
P3A ~~ P3B + P3C + P4A + P4B + P4C 
P3B ~~ P3C + P4A + P4B + P4C 
P3C ~~ P4A + P4B + P4C 
P4A ~~ P4B + P4C 
P4B ~~ P4C

SC_y1 =~ NA*Q11 + lambda1*Q11 + lambda2*Q10 + lambda3*Q8A + lambda4*Q8B + lambda5*Q8C + lambda6*Q9A + lambda7*Q9B + lambda8*Q9C

SC_y1 ~~ SC_y1
SC_y1 ~ 1

Q11 ~ int1*1 
Q10 ~ int2*1
Q8A ~ int3*1
Q8B ~ int4*1
Q8C ~ int5*1
Q9A ~ int6*1 
Q9B ~ int7*1
Q9C ~ int8*1

# Covariances between relational trust items
Q8A ~~ Q8B + Q8C + Q9A + Q9B + Q9C 
Q8B ~~ Q8C + Q9A + Q9B + Q9C 
Q8C ~~ Q9A + Q9B + Q9C 
Q9A ~~ Q9B + Q9C 
Q9B ~~ Q9C

SC_y2 =~ NA*R6 + lambda1*R6 + lambda2*R5 + lambda3*R3A + lambda4*R3B + lambda5*R3C + lambda6*R4A + lambda7*R4B + lambda8*R4C

SC_y2 ~~ SC_y2
SC_y2 ~ 1

R6 ~ int1*1
R5 ~ int2*1
R3A ~ int3*1
R3B ~ int4*1 
R3C ~ int5*1
R4A ~ int6*1
R4B ~ int7*1 
R4C ~ int8*1

# Covariances between relational trust items
R3A ~~ R3B + R3C + R4A + R4B + R4C 
R3B ~~ R3C + R4A + R4B + R4C 
R3C ~~ R4A + R4B + R4C 
R4A ~~ R4B + R4C 
R4B ~~ R4C

SC_y3 =~ NA*S12A + lambda1*S12A + lambda2*S11 + lambda3*S9A + lambda4*S9B + lambda5*S9C + lambda6*S10A + lambda7*S10B + lambda8*S10C

SC_y3 ~~ SC_y3
SC_y3 ~ 1

S12A ~ int1*1
S11 ~ int2*1
S9A ~ int3*1
S9B ~ int4*1 
S9C ~ int5*1
S10A ~ int6*1
S10B ~ int7*1 
S10C ~ int8*1

# Covariances between relational trust items
S9A ~~ S9B + S9C + S10A + S10B + S10C 
S9B ~~ S9C + S10A + S10B + S10C 
S9C ~~ S10A + S10B + S10C 
S10A ~~ S10B + S10C 
S10B ~~ S10C
'

covs <- '
P6 ~~ Q11 + R6 + S12A
Q11 ~~ R6 + S12A
R6 ~~ S12A

P5 ~~ Q10 + R5 + S11
Q10 ~~ R5 + S11
R5 ~~ S11

P3A ~~ Q8A + R3A + S9A
Q8A ~~ R3A + S9A
R3A ~~ S9A

P3B ~~ Q8B + R3B + S9B
Q8B ~~ R3B + S9B
R3B ~~ S9B

P3C ~~ Q8C + R3C + S9C
Q8C ~~ R3C + S9C
R3C ~~ S9C

P4A ~~ Q9A + R4A + S10A
Q9A ~~ R4A + S10A
R4A ~~ S10A

P4B ~~ Q9B + R4B + S10B
Q9B ~~ R4B + S10B
R4B ~~ S10B

P4C ~~ Q9C + R4C + S10C
Q9C ~~ R4C + S10C
R4C ~~ S10C

SC_pre ~~ SC_y1 + SC_y2 + SC_y3
SC_y1 ~~ SC_y2 + SC_y3
SC_y2 ~~ SC_y3
'
#fit_meas_y3 <- sem(model = meas_mod_y3, data = FASTData, estimator = "ML", missing = "ML")
#summary(fit_meas, fit.measures = T)

strong <- paste(strong_meas_mod, covs)
```

## Partially Strong Invariance Model
```{r, warning = F}
partial_strong_mod <- '
SC_pre =~ NA*P6 + lambda1*P6 + lambda2*P5 + lambda3*P3A + lambda4*P3B + lambda5*P3C + lambda6*P4A + lambda7*P4B + lambda8*P4C

SC_pre ~~ 1*SC_pre
SC_pre ~ 1

P6 ~ 1
P5 ~ 1
P3A ~ 1
P3B ~ 1 
P3C ~ 1
P4A ~ 1
P4B ~ int7*1 
P4C ~ int8*1

# Covariances between relational trust items
P3A ~~ P3B + P3C + P4A + P4B + P4C 
P3B ~~ P3C + P4A + P4B + P4C 
P3C ~~ P4A + P4B + P4C 
P4A ~~ P4B + P4C 
P4B ~~ P4C

SC_y1 =~ NA*Q11 + lambda1*Q11 + lambda2*Q10 + lambda3*Q8A + lambda4*Q8B + lambda5*Q8C + lambda6*Q9A + lambda7*Q9B + lambda8*Q9C

SC_y1 ~~ 1*SC_y1
SC_y1 ~ 1

Q8A ~ 1
Q8B ~ 1
Q8C ~ 1
Q9A ~ 1 
Q9B ~ 1
Q9C ~ 1
Q11 ~ int7*1 
Q10 ~ int8*1

# Covariances between relational trust items
Q8A ~~ Q8B + Q8C + Q9A + Q9B + Q9C 
Q8B ~~ Q8C + Q9A + Q9B + Q9C 
Q8C ~~ Q9A + Q9B + Q9C 
Q9A ~~ Q9B + Q9C 
Q9B ~~ Q9C

SC_y2 =~ NA*R6 + lambda1*R6 + lambda2*R5 + lambda3*R3A + lambda4*R3B + lambda5*R3C + lambda6*R4A + lambda7*R4B + lambda8*R4C

SC_y2 ~~ 1*SC_y2
SC_y2 ~ 1

R6 ~ 1
R5 ~ 1
R3A ~ 1
R3B ~ 1 
R3C ~ 1
R4A ~ 1
R4B ~ int7*1 
R4C ~ int8*1

# Covariances between relational trust items
R3A ~~ R3B + R3C + R4A + R4B + R4C 
R3B ~~ R3C + R4A + R4B + R4C 
R3C ~~ R4A + R4B + R4C 
R4A ~~ R4B + R4C 
R4B ~~ R4C

SC_y3 =~ NA*S12A + lambda1*S12A + lambda2*S11 + lambda3*S9A + lambda4*S9B + lambda5*S9C + lambda6*S10A + lambda7*S10B + lambda8*S10C

SC_y3 ~~ 1*SC_y3
SC_y3 ~ 1

S12A ~ 1
S11 ~ 1
S9A ~ 1
S9B ~ 1 
S9C ~ 1
S10A ~ 1
S10B ~ int7*1 
S10C ~ int8*1

# Covariances between relational trust items
S9A ~~ S9B + S9C + S10A + S10B + S10C 
S9B ~~ S9C + S10A + S10B + S10C 
S9C ~~ S10A + S10B + S10C 
S10A ~~ S10B + S10C 
S10B ~~ S10C
'

covs <- '
P6 ~~ Q11 + R6 + S12A
Q11 ~~ R6 + S12A
R6 ~~ S12A

P5 ~~ Q10 + R5 + S11
Q10 ~~ R5 + S11
R5 ~~ S11

P3A ~~ Q8A + R3A + S9A
Q8A ~~ R3A + S9A
R3A ~~ S9A

P3B ~~ Q8B + R3B + S9B
Q8B ~~ R3B + S9B
R3B ~~ S9B

P3C ~~ Q8C + R3C + S9C
Q8C ~~ R3C + S9C
R3C ~~ S9C

P4A ~~ Q9A + R4A + S10A
Q9A ~~ R4A + S10A
R4A ~~ S10A

P4B ~~ Q9B + R4B + S10B
Q9B ~~ R4B + S10B
R4B ~~ S10B

P4C ~~ Q9C + R4C + S10C
Q9C ~~ R4C + S10C
R4C ~~ S10C

SC_pre ~~ SC_y1 + SC_y2 + SC_y3
SC_y1 ~~ SC_y2 + SC_y3
SC_y2 ~~ SC_y3
'
#fit_meas_y3 <- sem(model = meas_mod_y3, data = FASTData, estimator = "ML", missing = "ML")
#summary(fit_meas, fit.measures = T)

partial <- paste(partial_strong_mod, covs)
```


## Partial Strong Invariance

```{r, warning = F, eval = F}
partial_strong_mod <- '
SC_pre =~ NA*P6 + lambda1*P6 + lambda2*P5 + lambda3*P3A + lambda4*P3B + lambda5*P3C + lambda6*P4A + lambda7*P4B + lambda8*P4C

SC_pre ~~ SC_pre
SC_pre ~ 1

P6 ~ NA*1
P5 ~ NA*1
P3A ~ NA*1
P3B ~ NA*1 
P3C ~ NA*1
P4A ~ NA*1
P4B ~ int7*1 
P4C ~ int8*1

0 == (int7 + int8)/2

# Covariances between relational trust items
P3A ~~ P3B + P3C + P4A + P4B + P4C 
P3B ~~ P3C + P4A + P4B + P4C 
P3C ~~ P4A + P4B + P4C 
P4A ~~ P4B + P4C 
P4B ~~ P4C

SC_y1 =~ NA*Q11 + lambda1*Q11 + lambda2*Q10 + lambda3*Q8A + lambda4*Q8B + lambda5*Q8C + lambda6*Q9A + lambda7*Q9B + lambda8*Q9C

SC_y1 ~~ SC_y1
SC_y1 ~ 1

Q11 ~ NA*1 
Q10 ~ NA*1
Q8A ~ NA*1
Q8B ~ NA*1
Q8C ~ NA*1
Q9A ~ NA*1 
Q9B ~ int7*1
Q9C ~ int8*1

# Covariances between relational trust items
Q8A ~~ Q8B + Q8C + Q9A + Q9B + Q9C 
Q8B ~~ Q8C + Q9A + Q9B + Q9C 
Q8C ~~ Q9A + Q9B + Q9C 
Q9A ~~ Q9B + Q9C 
Q9B ~~ Q9C

SC_y2 =~ NA*R6 + lambda1*R6 + lambda2*R5 + lambda3*R3A + lambda4*R3B + lambda5*R3C + lambda6*R4A + lambda7*R4B + lambda8*R4C

SC_y2 ~~ SC_y2
SC_y2 ~ 1

R6 ~ NA*1
R5 ~ NA*1
R3A ~ NA*1
R3B ~ NA*1 
R3C ~ NA*1
R4A ~ NA*1
R4B ~ int7*1 
R4C ~ int8*1

# Covariances between relational trust items
R3A ~~ R3B + R3C + R4A + R4B + R4C 
R3B ~~ R3C + R4A + R4B + R4C 
R3C ~~ R4A + R4B + R4C 
R4A ~~ R4B + R4C 
R4B ~~ R4C

SC_y3 =~ NA*S12A + lambda1*S12A + lambda2*S11 + lambda3*S9A + lambda4*S9B + lambda5*S9C + lambda6*S10A + lambda7*S10B + lambda8*S10C

SC_y3 ~~ SC_y3
SC_y3 ~ 1

S12A ~ NA*1
S11 ~ NA*1
S9A ~ NA*1
S9B ~ NA*1 
S9C ~ NA*1
S10A ~ NA*1
S10B ~ int7*1 
S10C ~ int8*1

# Covariances between relational trust items
S9A ~~ S9B + S9C + S10A + S10B + S10C 
S9B ~~ S9C + S10A + S10B + S10C 
S9C ~~ S10A + S10B + S10C 
S10A ~~ S10B + S10C 
S10B ~~ S10C
'

covs <- '
P6 ~~ Q11 + R6 + S12A
Q11 ~~ R6 + S12A
R6 ~~ S12A

P5 ~~ Q10 + R5 + S11
Q10 ~~ R5 + S11
R5 ~~ S11

P3A ~~ Q8A + R3A + S9A
Q8A ~~ R3A + S9A
R3A ~~ S9A

P3B ~~ Q8B + R3B + S9B
Q8B ~~ R3B + S9B
R3B ~~ S9B

P3C ~~ Q8C + R3C + S9C
Q8C ~~ R3C + S9C
R3C ~~ S9C

P4A ~~ Q9A + R4A + S10A
Q9A ~~ R4A + S10A
R4A ~~ S10A

P4B ~~ Q9B + R4B + S10B
Q9B ~~ R4B + S10B
R4B ~~ S10B

P4C ~~ Q9C + R4C + S10C
Q9C ~~ R4C + S10C
R4C ~~ S10C

SC_pre ~~ SC_y1 + SC_y2 + SC_y3
SC_y1 ~~ SC_y2 + SC_y3
SC_y2 ~~ SC_y3
'
#fit_meas_y3 <- sem(model = meas_mod_y3, data = FASTData, estimator = "ML", missing = "ML")
#summary(fit_meas, fit.measures = T)

strong <- paste(partial_strong_mod, covs)
fit_strong <- sem(model = strong, data = FASTData, estimator = "ML", missing = "ML")

summary(fit_strong)

models <- c("Configural" = fit_configural, "Weak" = fit_weak, "Strong" = fit_strong)

compareLavaan(models,
              fitmeas = c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"), 
              chidif = T,
              nesting = "Configural > Weak > Strong", 
              digits = 3)
```

```{r, warning = F, eval = F}
partial_strong_mod <- '
SC_pre =~ NA*P6 + lambda1*P6 + lambda2*P5 + lambda3*P3A + lambda4*P3B + lambda5*P3C + lambda6*P4A + lambda7*P4B + lambda8*P4C

SC_pre ~~ 1*SC_pre
SC_pre ~ NA*1

P6 ~ NA*1
P5 ~ NA*1
P3A ~ NA*1
P3B ~ NA*1 
P3C ~ NA*1
P4A ~ NA*1
P4B ~ NA*1 
P4C ~ int8*1

# Covariances between relational trust items
P3A ~~ P3B + P3C + P4A + P4B + P4C 
P3B ~~ P3C + P4A + P4B + P4C 
P3C ~~ P4A + P4B + P4C 
P4A ~~ P4B + P4C 
P4B ~~ P4C

SC_y1 =~ NA*Q11 + lambda1*Q11 + lambda2*Q10 + lambda3*Q8A + lambda4*Q8B + lambda5*Q8C + lambda6*Q9A + lambda7*Q9B + lambda8*Q9C

SC_y1 ~~ 1*SC_y1
SC_y1 ~ 1

Q11 ~ NA*1 
Q10 ~ NA*1
Q8A ~ NA*1
Q8B ~ NA*1
Q8C ~ NA*1
Q9A ~ NA*1 
Q9B ~ NA*1
Q9C ~ int8*1

# Covariances between relational trust items
Q8A ~~ Q8B + Q8C + Q9A + Q9B + Q9C 
Q8B ~~ Q8C + Q9A + Q9B + Q9C 
Q8C ~~ Q9A + Q9B + Q9C 
Q9A ~~ Q9B + Q9C 
Q9B ~~ Q9C

SC_y2 =~ NA*R6 + lambda1*R6 + lambda2*R5 + lambda3*R3A + lambda4*R3B + lambda5*R3C + lambda6*R4A + lambda7*R4B + lambda8*R4C

SC_y2 ~~ 1*SC_y2
SC_y2 ~ 1

R6 ~ NA*1
R5 ~ NA*1
R3A ~ NA*1
R3B ~ NA*1 
R3C ~ NA*1
R4A ~ NA*1
R4B ~ NA*1 
R4C ~ int8*1

# Covariances between relational trust items
R3A ~~ R3B + R3C + R4A + R4B + R4C 
R3B ~~ R3C + R4A + R4B + R4C 
R3C ~~ R4A + R4B + R4C 
R4A ~~ R4B + R4C 
R4B ~~ R4C

SC_y3 =~ NA*S12A + lambda1*S12A + lambda2*S11 + lambda3*S9A + lambda4*S9B + lambda5*S9C + lambda6*S10A + lambda7*S10B + lambda8*S10C

SC_y3 ~~ 1*SC_y3
SC_y3 ~ 1

S12A ~ NA*1
S11 ~ NA*1
S9A ~ NA*1
S9B ~ NA*1 
S9C ~ NA*1
S10A ~ NA*1
S10B ~ NA*1 
S10C ~ int8*1

# Covariances between relational trust items
S9A ~~ S9B + S9C + S10A + S10B + S10C 
S9B ~~ S9C + S10A + S10B + S10C 
S9C ~~ S10A + S10B + S10C 
S10A ~~ S10B + S10C 
S10B ~~ S10C
'

covs <- '
P6 ~~ Q11 + R6 + S12A
Q11 ~~ R6 + S12A
R6 ~~ S12A

P5 ~~ Q10 + R5 + S11
Q10 ~~ R5 + S11
R5 ~~ S11

P3A ~~ Q8A + R3A + S9A
Q8A ~~ R3A + S9A
R3A ~~ S9A

P3B ~~ Q8B + R3B + S9B
Q8B ~~ R3B + S9B
R3B ~~ S9B

P3C ~~ Q8C + R3C + S9C
Q8C ~~ R3C + S9C
R3C ~~ S9C

P4A ~~ Q9A + R4A + S10A
Q9A ~~ R4A + S10A
R4A ~~ S10A

P4B ~~ Q9B + R4B + S10B
Q9B ~~ R4B + S10B
R4B ~~ S10B

P4C ~~ Q9C + R4C + S10C
Q9C ~~ R4C + S10C
R4C ~~ S10C

SC_pre ~~ SC_y1 + SC_y2 + SC_y3
SC_y1 ~~ SC_y2 + SC_y3
SC_y2 ~~ SC_y3
'
#fit_meas_y3 <- sem(model = meas_mod_y3, data = FASTData, estimator = "ML", missing = "ML")
#summary(fit_meas, fit.measures = T)

strong <- paste(partial_strong_mod, covs)
fit_strong <- sem(model = strong, data = FASTData, estimator = "ML", missing = "ML")

summary(fit_strong)

models <- c("Configural" = fit_configural, "Weak" = fit_weak, "Strong" = fit_strong)

compareLavaan(models,
              fitmeas = c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"), 
              chidif = T,
              nesting = "Configural > Weak > Strong", 
              digits = 3)

# library(semTools)
# models <- list(fit.configural = fit_configural, fit.loadings = fit_weak, fit.intercepts = fit_strong)
# partialInvariance(fit = models, type = "scalar")
```

```{r, warning = F}
fit_strong <- sem(model = strong, data = FASTData, estimator = "ML", missing = "ML")
fitMeasures(fit_strong, fit.measures = c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
```

## Model Invariance Comparisons

```{r}
models <- c("Configural" = fit_configural, "Weak" = fit_weak, "Strong" = fit_strong)

compareLavaan(models,
              fitmeas = c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"), 
              chidif = T,
              nesting = "Configural > Weak > Strong", 
              digits = 3)
```

# Factor Scores

## Creating and Merging Factor Scores to Dataset

```{r, warning = F}
fit_pre <- sem(model = meas_mod_pre, data = FASTData, estimator = "ML", missing = "ML")
fit_y1 <- sem(model = meas_mod_y1, data = FASTData, estimator = "ML", missing = "ML")
fit_y2 <- sem(model = meas_mod_y2, data = FASTData, estimator = "ML", missing = "ML")
fit_y3 <- sem(model = meas_mod_y3, data = FASTData, estimator = "ML", missing = "ML")

fitPredict <- as.data.frame(predict(fit_pre))
FASTData <- cbind(FASTData, fitPredict)

fitPredict2 <- as.data.frame(predict(fit_y1))
FASTData <- cbind(FASTData, fitPredict2)

fitPredict3 <- as.data.frame(predict(fit_y2))
FASTData <- cbind(FASTData, fitPredict3)

fitPredict4 <- as.data.frame(predict(fit_y3))
FASTData <- cbind(FASTData, fitPredict4)
```

```{r, message = F, warning = F}
FASTData_SC_FAST <- filter(FASTData, FAST == "FAST Program")
FASTData_SC_FAST <- select(FASTData_SC_FAST, "SC_pre", "SC_y1", "SC_y2", "SC_y3")

FASTData_SC_Cntrl <- filter(FASTData, FAST == "Control School")
FASTData_SC_Cntrl <- select(FASTData_SC_Cntrl, "SC_pre", "SC_y1", "SC_y2", "SC_y3")

p1 <- FASTData_SC_FAST[,1:4] |> 
  pivot_longer(everything(), names_to = "vars") |>
  ggplot(aes(x = value, fill = vars)) +
  geom_density() + 
  scale_y_continuous(limits = c(0, 0.8)) +
  facet_wrap(~ vars, nrow = 4) +
  labs(x = " FAST Program Social Capital Factor Score", y = "Density", 
       title = "Density Plots of Social Capital Factor Scores,") +
  theme_minimal() +
  scale_fill_grey() +
  theme(legend.position = "none")

p2 <- FASTData_SC_Cntrl[,1:4] |> 
  pivot_longer(everything(), names_to = "vars") |>
  ggplot(aes(x = value, fill = vars)) +
  geom_density() + 
  scale_y_continuous(limits = c(0, 0.8)) +
  facet_wrap(~ vars, nrow = 4) +
  labs(x = "Control School Social Capital Factor Score", y = "Density", 
       title = "") +
  theme_minimal() +
  scale_fill_grey() +
  theme(axis.title.y = element_blank(), axis.text.y=element_blank(), legend.position = "none")

p1 + p2
```

## Exporting Updated Dataset

```{r}
write.csv(FASTData, "/Users/jayjeffries/Desktop/Longitudinal/SEM/Assignments/Final Project/FAST SEM Project/FASTData_FS.csv", row.names = T)
```
