---
title: "Family and Schools Together <br /> Structural Models"
author: "Jay Jeffries"
date: "11/14/2022"
output:
  rmdformats::robobook:
    code_folding: show
    self_contained: true
    thumbnails: false
    lightbox: true
pkgdown:
  as_is: true  
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = T)

library(pacman)

p_load("tidyverse", "lavaan", "skimr", "haven", "survey", "VIM", "semPlot", "tidySEM", "rmdformats", "semTable", 
       "kableExtra", "htmlTable", "vioplot", install = T)

FASTData <- read_csv("/Users/jayjeffries/Desktop/Longitudinal/SEM/Assignments/Final Project/FAST SEM Project/FASTData_FS.csv")

FASTData$PreTest <- FASTData$SC_pre
FASTData$Year1 <- FASTData$SC_y1
FASTData$Year2 <- FASTData$SC_y2
FASTData$Year3 <- FASTData$SC_y3

long_FASTData <- pivot_longer(
  FASTData,
  cols = PreTest:Year3,
  names_to = "Time",
  values_to = "SocialCapital",
)

write.csv(long_FASTData, "/Users/jayjeffries/Desktop/Longitudinal/SEM/Assignments/Final Project/FAST SEM Project/long_FASTData.csv", row.names = T)
```

# Visualization & Descriptives

```{r, message = F, echo = F, warning = F}
mean <- base::mean

ggplot(data = subset(long_FASTData, FAST == "FAST Program"), aes(x = Time, y = SocialCapital, group = STUDENTID)) +
  geom_point(color = "darkgray", size = 1, shape = 3) +
  stat_smooth(method = "lm", se = FALSE, color = "lightgray", size = .1) +
  geom_jitter(color = "darkgray", size = .5, shape = 3) +
  stat_summary(aes(group = 1, color = "Timepoint Average"), geom = "point", fun = mean, shape = 17, size = 5) +
  scale_color_manual(values = "#FFA500") + 
  scale_y_continuous(limits = c(1, 4), n.breaks = 4) +
  geom_smooth(aes(group = 1), method = lm, se = F, color = "black", size = .7) +
  labs(colour = NULL, title = "Plot of Student Social Capital Across Time, FAST Intervention Group",
       x = "Timepoint", y = "Social Capital Factor Score", 
       tag = "Figure 1.") +
  theme_classic() + 
  theme(axis.title = element_text(face = "bold"), 
        plot.title = element_text(face = "bold"))

ggplot(data = subset(long_FASTData, FAST == "Control School"), aes(x = Time, y = SocialCapital, group = STUDENTID)) +
  geom_point(color = "darkgray", size = 1, shape = 3) +
  stat_smooth(method = "lm", se = FALSE, color = "lightgray", size = .1) +
  geom_jitter(color = "darkgray", size = .5, shape = 3) +
  stat_summary(aes(group = 1, color = "Timepoint Average"), geom = "point", fun = mean, shape = 17, size = 5) +
  scale_color_manual(values = "#FFA500") + 
  scale_y_continuous(limits = c(1, 4), n.breaks = 4) +
  geom_smooth(aes(group = 1), method = lm, se = F, color = "black", size = .7) +
  labs(colour = NULL, title = "Plot of Student Social Capital Across Time, Control Schools",
       x = "Timepoint", y = "Social Capital Factor Score", 
       tag = "Figure 2.") +
  theme_classic() + 
  theme(axis.title = element_text(face = "bold"), 
        plot.title = element_text(face = "bold"))

DescTable = FASTData %>%
  summarise( 
    "Social Capital Pre-Test" = round(mean(SC_pre, na.rm = T), 3),
    "Social Capital Year 1 \n Post-Test" = round(mean(SC_y1, na.rm = T), 3),
    "Social Capital Year 2 \n Post-Test" = round(mean(SC_y2, na.rm = T), 3),
    "Social Capital Year 3 \n Post-Test" = round(mean(SC_y3, na.rm = T), 3),
  )

kable(caption = "Table 1. Social Capital Descriptives", align = 'c', x = DescTable) |>
  add_header_above(c("Mean Social Capital Factor Score Across Time" = 4)) |>
  kable_paper("hover", font_size = 13, full_width = T) |>
  kable_styling(html_font = "serif", row_label_position = "c")

# Trellis plot
long_FASTData1 <- long_FASTData |>
  filter(FAST == "FAST Program")

long_FASTData1 <- long_FASTData1 |> drop_na(SocialCapital)

long_FASTData1 <- long_FASTData1 |> 
  group_by(STUDENTID) |>
  filter(n() == 4) 

set.seed(13130)
stu_IDs <- sample(long_FASTData1$STUDENTID, 20, replace = T)

FAST_random <- long_FASTData1 |> filter(STUDENTID %in% stu_IDs)

FAST_random$Timepoint <- ifelse(FAST_random$Time == "PreTest", 0, 
                                ifelse(FAST_random$Time == "Year1", 1,
                                       ifelse(FAST_random$Time == "Year2", 2,
                                              ifelse(FAST_random$Time == "Year3", 3, NA))))

lattice <- ggplot(FAST_random, aes(x = Timepoint, y = SocialCapital)) + 
  geom_point(shape = 1) + 
  geom_smooth(method = "lm", color = "orange", se = F) + 
  facet_wrap(~ STUDENTID, nrow = 4, ncol = 5) + 
  ylim(0, 4) +
  ggtitle("Lattice Plot of Student Social Capital Scores \n for n = 20 Random FAST Students") +
  theme_minimal() + 
  theme(axis.title = element_text(face = "bold"), 
        plot.title = element_text(face = "bold")) +
  theme(text = element_text(family = "Times New Roman"))
lattice

FASTData_FAST <- FASTData |>
  filter(FAST == "FAST Program")

vioplot_FAST <- vioplot(FASTData_FAST$SC_pre, FASTData_FAST$SC_y1, FASTData_FAST$SC_y2, FASTData_FAST$SC_y2,
        col = "white", names=c("Pre-Test", "Year 1", "Year 2", "Year 3"), ylim = c(1, 4),
        main = "Social Capital Score by Timepoints Violin Plots", xlab= "Timepoint", ylab= "Social Capital")
vioplot_FAST

FASTData_FAST <- long_FASTData |>
  filter(FAST == "FAST Program")

ggplot(long_FASTData, aes(x = Time, y = SocialCapital, fill = FAST)) + 
  geom_violin(aes(color = FAST), trim = FALSE, position = position_dodge(0.9)) +
  geom_boxplot(aes(color = FAST), width = 0.15, position = position_dodge(0.9)) +
  scale_fill_manual(values = c("#66545e", "#a39193")) +
  scale_color_manual(values = c("gray", "gray")) +
  ggtitle("Violin Plot of Student Social Capital Scores") +
  theme_minimal() + 
  theme(axis.title = element_text(face = "bold", size = 12), 
        plot.title = element_text(face = "bold", size = 12)) +
  theme(text = element_text(family = "Times New Roman"))
```

# Modelling

```{r, warning = F}
pre_SC_lat <- '
pre_SC =~ 1*P6 + P5 + P3A + P3B + P3C + P4A + P4B + P4C
pre_SC ~~ NA*pre_SC
pre_SC ~ 0*1

P3A ~~ P3B + P3C + P4A + P4B + P4C
P3B ~~ P3C + P4A + P4B + P4C
P3C ~~ P4A + P4B + P4C
P4A ~~ P4B + P4C
P4B ~~ P4C
'

y1_SC_lat <- '
y1_SC =~ 1*Q11 + Q10 + Q8A + Q8B + Q8C + Q9A + Q9B + Q9C
y1_SC ~~ NA*y1_SC
y1_SC ~ 0*1

Q8A ~~ Q8B + Q8C + Q9A + Q9B + Q9C
Q8B ~~ Q8C + Q9A + Q9B + Q9C
Q8C ~~ Q9A + Q9B + Q9C
Q9A ~~ Q9B + Q9C
Q9B ~~ Q9C
'

y2_SC_lat <- '
y2_SC =~ 1*R6 + R5 + R3A + R3B + R3C + R4A + R4B + R4C
y2_SC ~~ NA*y2_SC
y2_SC ~ 0*1

R3A ~~ R3B + R3C + R4A + R4B + R4C
R3B ~~ R3C + R4A + R4B + R4C
R3C ~~ R4A + R4B + R4C
R4A ~~ R4B + R4C
R4B ~~ R4C
'

y3_SC_lat <- '
y3_SC =~ 1*S12A + S11 + S9A + S9B + S9C + S10A + S10B + S10C
y3_SC ~~ NA*y3_SC
y3_SC ~ 0*1

S9A ~~ S9B + S9C + S10A + S10B + S10C
S9B ~~ S9C + S10A + S10B + S10C
S9C ~~ S10A + S10B + S10C
S10A ~~ S10B + S10C
S10B ~~ S10C
'

covs <- '
pre_SC ~~ y1_SC + y2_SC + y3_SC
y1_SC ~~ y2_SC + y3_SC
y2_SC ~~ y3_SC
'

meas_mods <- paste(pre_SC_lat, y1_SC_lat, y2_SC_lat, y3_SC_lat, covs)
fit_LCS <- sem(model = meas_mods, data = FASTData, estimator = "MLR", missing = "ML")
summary(fit_LCS, fit.measures = T)
```

## Linear Growth Model

```{r, warning = F}
model_int <- '
int =~ 1*SC_pre + 1*SC_y1 + 1*SC_y2 + 1*SC_y3
'

model_slp <- '
slp =~ 0*SC_pre + 1*SC_y1 + 2*SC_y2 + 3*SC_y3
'

linear_model <- paste(model_int, model_slp)

fit_linear <- growth(model = linear_model, data = FASTData, estimator = "MLR", missing = "ML")

fitMeasures(fit_linear, fit.measures = c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))

summary(fit_linear, fit.measures = T)
```

## Shape and Form Growth Model

```{r, warning = F}
model_form <- '
slp =~ 0*SC_pre + NA*SC_y1 + NA*SC_y2 + 1*SC_y3
'

shapeform_model <- paste(model_int, model_form)

fit_shapeform <- growth(model = shapeform_model, data = FASTData, estimator = "MLR", missing = "ML")
fitMeasures(fit_shapeform, fit.measures = c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
```

## Quadratic Growth Model

```{r, warning = F}
model_int <- '
int =~ 1*SC_pre + 1*SC_y1 + 1*SC_y2 + 1*SC_y3
'

model_slp <- '
slp =~ 0*SC_pre + 1*SC_y1 + 2*SC_y2 + 3*SC_y3
'

model_quad <- '
quad =~ 0*SC_pre + 1*SC_y1 + 4*SC_y2 + 9*SC_y3
'

quad_model <- paste(model_int, model_slp, model_quad)

fit_quad <- growth(model = quad_model, data = FASTData, estimator = "MLR", missing = "ML")
fitMeasures(fit_quad, fit.measures = c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
```

## Model Comparison

```{r, echo = F}
mods <- c("Linear" = fit_linear, "Quadratic" = fit_quad, "ShapeForm" = fit_shapeform)

compareLavaan(mods,
              fitmeas = c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"), 
              chidif = T,
              nesting = "Quadratic > Linear > ShapeForm", 
              digits = 3)
```

## Linear Growth Curve Structural Model

```{r, warning = F}
model_int <- '
int =~ 1*SC_pre + 1*SC_y1 + 1*SC_y2 + 1*SC_y3
'

model_slp <- '
slp =~ 0*SC_pre + 1*SC_y1 + 2*SC_y2 + 3*SC_y3
'

structural <- '
slp ~ LUNCH_STATUS + S2B + FAST + ATTENDANCE_PCT + PHQ2 + C3YRSTCH 
int ~ LUNCH_STATUS + S2B + FAST + ATTENDANCE_PCT + PHQ2 + C3YRSTCH 
'

linear_model <- paste(model_int, model_slp, structural)

final_linear <- growth(model = linear_model, data = FASTData, estimator = "MLR", missing = "ML")

summary(final_linear, fit.measures = T)
```

### Unstandardized Solution
```{r}
summary(final_linear, standardized = T, fit.measures = T, ci = T)
```

### Standardized Solution
```{r}
standardizedsolution(final_linear, level=.95)
```

### Linear Growth Curve Structural Model Figure
```{r}
prepare_graph(
  edges = get_edges(final_linear) |> filter(!str_detect(label_results, "Variances")),
  nodes = get_nodes(final_linear) |> mutate(label = name),
  layout = get_layout("LUNCH_STATUS", "S2B", "FAST", "ATTENDANCE_PCT", "PHQ2", "C3YRSTCH", 
                      "", "int", "", "", "slp", "",
                      "SC_pre", "SC_y1", "", "SC_y2", "", "SC_y3",
                       rows = 3), text_size = 2.8) |> plot()
```

# Final Model

## Linear Change Score Structural Model

```{r, warning = F, include = F}
SC_LCS <- '
# freed residuals
SC_pre ~~ SC_pre
SC_y1 ~~ SC_y1
SC_y2 ~~ SC_y2
SC_y3 ~~ SC_y3
 
# fix item means to zero
SC_pre ~ 0*1
SC_y1 ~ 0*1
SC_y2 ~ 0*1
SC_y3 ~ 0*1

# create phantom variables with loadings of 1
phant_pre =~ 1*SC_pre
phant_y1 =~ 1*SC_y1
phant_y2 =~ 1*SC_y2
phant_y3 =~ 1*SC_y3

# fix variance of phantom variables to 0
phant_pre ~~ 0*phant_pre
phant_y1 ~~ 0*phant_y1
phant_y2 ~~ 0*phant_y2
phant_y3 ~~ 0*phant_y3

# fix means of phantom variables to 0
phant_pre ~ 0*1
phant_y1 ~ 0*1
phant_y2 ~ 0*1
phant_y3 ~ 0*1

# to isolate changes, set autoregressions between phantom variables to 1
phant_y1 ~ 1*phant_pre
phant_y2 ~ 1*phant_y1
phant_y3 ~ 1*phant_y2

# to represent the change, create latent change scores with a loading of 1
change_prey1 =~ 1*phant_y1
change_y1y2 =~ 1*phant_y2
change_y2y3 =~ 1*phant_y3

# set variance of change scores to zero
change_prey1 ~~ 0*change_prey1
change_y1y2 ~~ 0*change_y1y2
change_y2y3 ~~ 0*change_y2y3

# set means of change scores to zero
change_prey1 ~ 0*1
change_y1y2 ~ 0*1
change_y2y3 ~ 0*1

# set the intercept of the trajectory based on the first timepoint
int =~ 1*phant_pre

# set the slope of the trajectory to indicate the same change between each interval
slp =~ 1*change_prey1 + 1*change_y1y2 + 1*change_y2y3

# estimate variance of trajectory
int ~~ int
slp ~~ slp

# estimate mean of the trajectory
int ~ 1
slp ~ 1

# estimate the covariance among parts of the trajectory
int ~~ slp
'

struct <- '
change_prey1 ~ C3YRSTCH + LUNCH_STATUS + S2B + FAST + ATTENDANCE_PCT + PHQ2
change_y1y2 ~ C3YRSTCH + LUNCH_STATUS + S2B + FAST + ATTENDANCE_PCT + PHQ2
change_y2y3 ~ C3YRSTCH + LUNCH_STATUS + S2B + FAST + ATTENDANCE_PCT + PHQ2
'

full_mod <- paste(SC_LCS, struct)

fit_LCS <- sem(model = full_mod, data = FASTData, estimator = "MLR", missing = "ML")

summary(fit_LCS, fit.measures = T)
```

```{r}
SC_LCS <- '
# freed variances
# pre_SC ~~ NA*pre_SC
# y1_SC ~~ NA*y1_SC
# y2_SC ~~ NA*y2_SC
# y3_SC ~~ NA*y3_SC
 
# fix latent means to zero
# pre_SC ~ 0*1
# y1_SC ~ 0*1
# y2_SC ~ 0*1
# y3_SC ~ 0*1

# create phantom variables with loadings of 1
phant_pre =~ 1*pre_SC
phant_y1 =~ 1*y1_SC
phant_y2 =~ 1*y2_SC
phant_y3 =~ 1*y3_SC

# fix variance of phantom variables to 0
phant_pre ~~ 0*phant_pre
phant_y1 ~~ 0*phant_y1
phant_y2 ~~ 0*phant_y2
phant_y3 ~~ 0*phant_y3

# fix means of phantom variables to 0
phant_pre ~ 0*1
phant_y1 ~ 0*1
phant_y2 ~ 0*1
phant_y3 ~ 0*1

# to isolate changes, set autoregressions between phantom variables to 1
phant_y1 ~ 1*phant_pre
phant_y2 ~ 1*phant_y1
phant_y3 ~ 1*phant_y2

# to represent the change, create latent change scores with a loading of 1
change_prey1 =~ 1*phant_y1
change_y1y2 =~ 1*phant_y2
change_y2y3 =~ 1*phant_y3

# set variance of change scores to zero
change_prey1 ~~ 0*change_prey1
change_y1y2 ~~ 0*change_y1y2
change_y2y3 ~~ 0*change_y2y3

# set means of change scores to zero
change_prey1 ~ 0*1
change_y1y2 ~ 0*1
change_y2y3 ~ 0*1

# set the intercept of the trajectory based on the first timepoint
int =~ 1*phant_pre

# set the slope of the trajectory to indicate the same change between each interval
slp =~ 1*change_prey1 + 1*change_y1y2 + 1*change_y2y3

# estimate variance of trajectory
int ~~ NA*int
slp ~~ NA*slp

# estimate mean of the trajectory
int ~ 1
slp ~ 1

# estimate the covariance among parts of the trajectory
int ~~ slp
'

covs <- '
pre_SC ~~ y1_SC + y2_SC + y3_SC
y1_SC ~~ y2_SC + y3_SC
y2_SC ~~ y3_SC
'

struct <- '
change_prey1 ~ C3YRSTCH + LUNCH_STATUS + S2B + FAST + ATTENDANCE_PCT + PHQ2 + phant_pre
change_y1y2 ~ C3YRSTCH + LUNCH_STATUS + S2B + FAST + ATTENDANCE_PCT + PHQ2 + phant_y1
change_y2y3 ~ C3YRSTCH + LUNCH_STATUS + S2B + FAST + ATTENDANCE_PCT + PHQ2 + phant_y2
'

full_mod <- paste(pre_SC_lat, y1_SC_lat, y2_SC_lat, y3_SC_lat, SC_LCS, covs, struct)
  
sem(model = full_mod, data = FASTData, estimator = "MLR", missing = "ML")

fit_LCS <- growth(model = full_mod, data = FASTData, estimator = "MLR", missing = "ML")

summary(fit_LCS, fit.measures = T)

std_LCS_output <- standardizedSolution(fit_LCS)
unstd_LCS_output <- parameterEstimates(fit_LCS)

write.csv(std_LCS_output, "/Users/jayjeffries/Desktop/Longitudinal/SEM/Assignments/Final Project/FAST SEM Project/Standardized_LCS_Output.csv")
write.csv(std_LCS_output, "/Users/jayjeffries/Desktop/Longitudinal/SEM/Assignments/Final Project/FAST SEM Project/Unstandardized_LCS_Output.csv")
```

```{r, eval = F, include = F}
# Not evaluated because these results are in the below table
summary(fit_LCS, standardized = T, fit.measures = T, ci = T)
```

### Standardized Solution
```{r}
standardizedsolution(fit_LCS, level=.95)
# Unstandardized solution found below
```

## Final Model Results

### Unstandardized Solution Table
```{r}
semTable(fit_LCS, columns = c("est", "se", "z", "p"), type = "html", print.results = F, 
         fits = c("npar", "chisq", "df", "pvalue", "chisq.scaled", "df.scaled", "pvalue.scaled", "cfi", 
                  "rmsea", "srmr")) |>
  htmlTable()
```

## Model Figure

### Linear Change Score Structural Model Figure
```{r}
prepare_graph(
  edges = get_edges(fit_LCS) |> filter(!str_detect(label_results, "Variances")),
  nodes = get_nodes(fit_LCS) |> mutate(label = name),
  layout = get_layout("int", "", "slp", "", "S2B", "C3YRSTCH", "FAST", "ATTENDANCE_PCT", "", "",  "",
                      "", "", "", "", "", "", "", "", "", "LUNCH_STATUS", "PHQ2",
                      "", "", "", "", "", "", "", "", "", "", "",
                      "", "", "change_prey1", "", "", "change_y1y2", "", "", "change_y2y3", "", "",
                      "phant_pre", "", "phant_y1", "", "", "phant_y2", "", "", "phant_y3", "", "",
                      "SC_pre", "", "SC_y1", "", "", "SC_y2", "", "", "SC_y3", "",  "",
                       rows = 6), spacing_y = 2, spacing_x = 1.5, text_size = 1.4, angle = 170) |> plot()
```


```{r, eval = F, include = F}
prepare_graph(
  edges = get_edges(fit_LCS) |> filter(!str_detect(label_results, "Variances")),
  nodes = get_nodes(fit_LCS) |> mutate(label = name),
  layout = get_layout(
                      "int", "", "FAST", "slp", "", "", "",
                      "", "LUNCH_STATUS", "", "", "PHQ2", "C3YRSTCH", "ATTENDANCE_PCT",
                      "", "", "", "", "", "", "S2B",
                      "", "", "change_prey1", "change_y1y2", "change_y2y3", "", "",
                      "phant_pre", "", "phant_y1", "phant_y2", "phant_y3", "", "",
                      "Soc_Cap_pre", "", "Soc_Cap_y1", "Soc_Cap_y2", "Soc_Cap_y3", "", "",
                       rows = 6), spacing_y = 2, spacing_x = 2, text_size = 1.55) |>
edit_graph({ color = "black" }) |> 
  if_edit(pval < .05, {color = "blue"}) |>
  plot()


graph <- prepare_graph(
  edges = get_edges(fit_LCS) |> filter(!str_detect(label_results, "Variances")),
  nodes = get_nodes(fit_LCS) |> mutate(label = name),
  layout = get_layout(
                      "int", "", "FAST", "slp", "", "", "",
                      "", "LUNCH_STATUS", "", "", "PHQ2", "C3YRSTCH", "ATTENDANCE_PCT",
                      "", "", "", "", "", "", "S2B",
                      "", "", "change_prey1", "change_y1y2", "change_y2y3", "", "",
                      "phant_pre", "", "phant_y1", "phant_y2", "phant_y3", "", "",
                      "Soc_Cap_pre", "", "Soc_Cap_y1", "Soc_Cap_y2", "Soc_Cap_y3", "", "",
                       rows = 6), spacing_y = 2, spacing_x = 2, text_size = 1.55)
nodes(graph)
```

